### 2022-11-24 paper/paper.qmd
Initialize

---
title: "Proof of concept Zstandard custom compression dictionary for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  ## Background

  Zstandard (Zstd) represents a lossless data compression algorithm that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a "training mode" pipeline, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements.

  ## Results

  ## Conclusions
bibliography: bibliography.bib 
---

# Introduction

The storage of biological data has represented a significant topic of research, with a number of challenges presented over subsequent generations of technological development, with current challenges discovered as the volume and complexity of data continues to increase[@d2018high;@li2014big].

# Materials and methods

# Results

# Discussion

# Conclusion

# Supporting information

::: {.supp}
## S1 Fig. {#s1-fig}

Bold the title sentence.

Add descriptive text after the title of the item (optional).
:::

::: {.supp}
## S2 Fig. {#s2-fig}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 File. {#s1-file}

Lorem ipsum.

:::

::: {.supp}

## S1 Video. {#s1-video}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 Appendix. {#s1-appendix}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

::: {.supp}
## S1 Table. {#s1-table}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

# Acknowledgments

I would like to express my sincere gratitude to Professor David Walsh for his advice that helped shape the research as it progressed; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; The Quarto developers for producing the Quarto publishing system along with the PLOS template used in this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments; Maximiliano Sandoval and contributors to the Citations app  for managing the paper's bibliography, and all other persons that had indirectly assisted through their programs and research.