### 2022-11-28 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

\PassOptionsToPackage{table}{xcolor}

\documentclass[
  10pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% marvosym package for additional characters
\usepackage{marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% Using natbib instead
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Exploring Zstandard user-provided dictionary compression for FASTA files},
  pdfauthor={Michael Persico},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Exploring Zstandard user-provided dictionary compression
for FASTA
files} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Michael Persico\textsuperscript{1}
\\
\bigskip
\textbf{1} Department of Biology, Concordia
University, Montreal, Quebec, Canada, 
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments
sections

% Use the asterisk to denote corresponding authorship and provide email address in note below.

\end{flushleft}

\section*{Abstract}
\hypertarget{background}{%
\subsection{Background}\label{background}}

Zstandard (Zstd) represents a universal, lossless data compression
standard and implementation that is highly configurable and is aimed at
coupling high compression ratios with fast compression/decompression
performance. Previous studies have paired specific Zstd configurations
with various file formats in bioinformatics to reduce total data volume.
This paper presents a ``training mode'' pipeline, written in the Julia
programming language, wherein a custom compression dictionary is
generated from a sample FASTA set in order to explore further
compression improvements and compare them to the compression performance
of Xz, Zlib, Bzip2, and Lz4 universal compressors.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\linenumbers\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The explosion of biological data has represented a significant topic of
research, with a number of challenges presented over subsequent
generations of technological development in regards to the management of
the increasing volume and complexity of
data\citep{d2018high, li2014big}. In response, emerging trends in data
management have lead to the development of novel, scalable methods for
the efficient transmission and storage of large amounts of
data\citep{sais2022intelligent}. With a potentially exponential quantity
of files, datasets, and other data resources to be handled, data
compression represents a method for reducing overall resource size by
encoding the original data into a compact form, thus helping to ease
storage requirements \citep{jayasankar2021survey}. Research into data
compression in the context of biological data began to pick up near the
turn of the 21st century as universal compression algorithms at the time
were not considered ideal for compressing DNA or RNA sequence data well,
which led to the introduction of purpose-built algorithms that addressed
the unique peculiarities of genomic data\citep{grumbach1994new}. At the
same time, new file formats were introduced, either text-based or
binary-based, for more accurate structuring and representation of
biological data, complementing new software
tools\citep{lipman1985rapid, mills2014common}.

The FASTA file format is a legacy of the original FASTA program for
finding sequence similarities with a query
sequence\citep{lipman1985rapid}. Each file can possess multiple
sequences, each paired with a description line distinguished by a
``\textgreater{}'' symbol followed by arbitrary text, usually a name
and/or summary description, on the same line. It is a commonly supported
file format in bioinformatics and has been the target for optimized data
compressors with competing claims for performance. The DELIMINATE
lossless algorithm was first proposed in 2012, wherein header and
sequence data are separated into DELIM-1 and DELIM-2 variants and a
two-phase process is pursued involving delta encoding, progressive
elimination of nucleotide characters, and 7-Zip
archiving\citep{mohammed2012deliminate}. The claims of better
compression/decompression performance of FASTA files were soon rivaled
by the introduction of the MFCompress tool, again separating headers and
sequence data but instead relying on probabilistic models to encode the
data\citep{pinho2014mfcompress}, which was then countered by the
Nucleotide Archival Format, a novel file format noteworthy in this
context for the inclusion of a Zstandard compression
step\citep{kryukov2019nucleotide}.

IETF RFC 8878, introduced by engineers at Facebook, defines Zstandard as
a lossless data compression/decompression
format\citep{collet2021rfc8478}. It is often abbreviated as ``Zstd'',
though such can also refer to Facebook's own implementation of the
algorithm written mostly in C\citep{facebook}. Content is sliced and
packaged into ``frames'' that are independent of one another defined as
either compressed data Zstandard frames or Skippable frames containing
custom user metadata\citep{collet2021rfc8478}. Zstd's backbone is the
use of Finite State Entropy and Huffman entropy encoding schemes that
replace data with coded forms independently of the
medium\citep{ezhilarasan_thambidurai_praveena_srinivasan_sumathi_2007, LU20171},
with the former compressing all symbols, though header information is
first encoded by the latter\citep{collet2021rfc8478}. Zstd, for small
data compression improvements, also functions as a dictionary coder,
meaning that although the algorithm is universal in the sense of being
applicable to a number of both text-based and binary-based data files,
it can be optimized for compacting characteristic data by ``training''
Zstd with a collection of sample files to build a set of common patterns
that allow for substitutions when compressing/decompressing, allowing
for further gains for similar data\citep{8316293}.

Common bioinformatics file formats often include a set structure with
the express purpose of representing specific kinds of biological data
composed of repeating elements, as is the case with FASTA files with
either nucleotide or amino acid sequences. In this work is described a
pipeline for building Zstd dictionaries via FASTA datasets along with a
comparison of Zstd compression/decompression performance with that of
several alternative lossless compressors for select datasets.

\hypertarget{materials-and-methods}{%
\section{Materials and methods}\label{materials-and-methods}}

A list of all direct and indirect dependencies can be found in the
repository's Manifest.toml file.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

\paragraph*{S1 Fig.}
\label{s1-fig}
{\textbf{Bold the title sentence.}} Add descriptive text after the title
of the item (optional).

\paragraph*{S2 Fig.}
\label{s2-fig}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 File.}
\label{s1-file}
{\textbf{Lorem ipsum.}}

\paragraph*{S1 Video.}
\label{s1-video}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{s1-appendix}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\paragraph*{S1 Table.}
\label{s1-table}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

I would like to express my sincere gratitude to Professor David Walsh
for his advice that helped shape the research as it progressed; My
friends and family that supported me throughout my studies; The Julia
community for their support before and throughout the writing of this
paper; The Quarto developers for producing the Quarto publishing
system\citep{Allaire_Quarto_2022} along with the PLOS template used for
this paper; Luca Di Maio and contributors to the Distrobox tool for the
ease of setting up the containerized development
environments\citep{maio}; Maximiliano Sandoval and contributors to the
Citations app for managing the paper's bibliography\citep{sandoval}, and
all other persons that had indirectly assisted through their programs
and research.

\nolinenumbers
  \bibliography{bibliography.bib}

\end{document}